Training a Gaming Agent on Brainwaves

### 2020-12-03 final/ErrP.tex
Subo la versión final del paper aceptado.

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
\IEEEPARstart{T}{he} effectiveness of today's human–machine interaction and artificial intelligence is limited by a communication bottleneck, as humans are required to translate high-level concepts into a machine-mandated sequence of instructions~\cite{Xu2020,CURSOR-CONTROL-PAPER}.   Hence, new interaction methods are required to increase the communication bandwidth between computers and humans or to produce alternative communications systems to increase the efficiency of this channel.  In this respect, video games have been widely used as test tools to assess new means of interactions~\cite{Carter2014,Barr2007}. Video gaming agents are computer programs that can sense the computer game environment, process information, and react accordingly within the environment.  They are used in the context of testing and evaluating artificial intelligence algorithms that aim to win the game or to behave like a real user player~\cite{Zhao2020}.
In this work, the feedback obtained from an observational human critic (OHC) in the form of electroencephalographic (EEG) signals is used to evaluate the operational performance of a gaming agent.  Observational human critics are silent subjects observing a computer gaming agent playing the game.

% Reescribí la primera oración porque el sujeto no tenía sentido con el predicado.  TODO
The feasibility of a distinct non-biological communication channel between the Central Nervous System (CNS) and a computer device has been previously proven with Brain Computer Interfaces (BCI) or Brain Machine Interfaces (BMI).~\cite{Vasiljevic2020}.  BCI systems provide a new input modality that can be used in computer game~\cite{Scherer2012,Nijholt2007}. This advancement is relevant in the context of the accessibility for video games~\cite{Aguado-Delgado2020} and the growing area of e-sports~\cite{Yakovlev2020}.

@@

Given the scenario, Reinforcement Learning (RL)~\cite{Sutton2018} stands out as a natural method to train the agent.  Reinforcement Learning refers to an algorithmic learning strategy inspired on how biological agents learn by exploring their environment while getting negative or positive feedback rewards.  The method aims to maximize positive rewards while minimizing negative feedback.  Thus, the learning problem is posed as a stochastic optimization strategy~\cite{Santos1999}.  Recently, this technique has been used extensively in the context of advances in artificial intelligence~\cite{Nguyen2020}. The influence of DeepBrain's AlphaGo project cannot be neglected, since it was the first to reach a very high proficiency when it won the complex game Go against several world champions~\cite{ALPHA-GO}.

@@

This work unfold as follows: in Section \ref{section:materials} the general layout of the cognitive game is described. Sections \ref{brainwavesession} and \ref{cognitive_experiment_details} outline the cognitive game procedure used to obtain rewards in the form of ErrP components. Section \ref{q_learning_step_alg} describes the gaming agent learning procedure.  Lastly, results and conclusions are exposed in Sections \ref{results} and \ref{conclusions}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{Images/complete_flow.png}
    \caption{Overview of the experimental procedure. Brainwaves are obtained by the OpenVibe Acquisition Server.  The Game Manager is responsible for generating the game screen, the game mechanics, and the game movements performed by the gaming agent.  It is also connected to the Acquisition Server to send stimulus information.  The captured information is stored by the OpenVibe Designer.  Offline, EEG signals are classified and they are linked to each game movement calculated by the Game Manager to determine proper rewards for each action.  This information is used by a Reinforcement Learning algorithm that iteratively trains a Q-Table in order to improve the performance of the agent that plays the game.}
    \label{diag:complete_flow}
\end{figure*}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.2]{Images/grid_initial_state.png}
\includegraphics[scale=0.2]{Images/grid_end_state.png}
\end{subfigure}
\centering
\caption{Grid system representation used in the cognitive game. The blue spot represents the initial location while the green spot represents the target location. Once the agent reaches the target spot, its color turns red to indicate the end of the play.}
\label{fig:game_representation}
\end{figure}

@@

This phase is divided into a sequence of run sessions and gaming agent training matches.  During the run session, the agent plays 200 matches guided by a specific Q-Table with a 5\% chance of randomly selecting a movement, to reduce deadlocks and loops.  Following the run session, the agent performs a single gaming agent training match.  The gaming agent starts first with a Q-Table initialized with zeros, so the initial policy for the agent is randomized.  For the agent to learn from the feedback generated by the OHC, movement actions are determined by the reply of the agent's actions that were taken during one brainwave session match, in an offline reinforcement learning scheme~\cite{Levine2020}. This allows the Q-Table to be learned based on the OHC's feedback from the movements the agent took, which were executed pseudo-randomly during the brainwave session.  The previously mentioned feedback is not explicit as it comes from the interpreted brain signal data. This implies that the reward is determined by the OHC's brain activity.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_closer.eps}
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_further.eps}
        \includegraphics[scale=1.0]{revisedimages/electrodes.png}
    \caption{Grand average of 2-seconds time-locked segments for all OHCs for the "move closer" (top) and "move further" (middle) condition. Zero corresponds to the onset of the stimulus, i.e. when the gaming agent moves. The increased height of the first peak around 0.4 seconds reflects the ErrP response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:grandaverage}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth,height=70pt]{revisedimages/cross_val.eps}
    %\includegraphics[width=290pt,height=90pt]{revisedimages/avg_5_segments.eps}
    \caption{Binary single trial classification score using five different classifiers while recognizing ErrP potentials for the eight OHCs.   Chance Level is 0.5.}
    \label{fig:classifiers}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.27]{Images/Average_steps/a.png}
\includegraphics[scale=0.27]{Images/Average_steps/b.png}
\centering
\includegraphics[scale=0.27]{Images/Average_steps/c.png}
 \includegraphics[scale=0.27]{Images/Average_steps/d.png}
 \centering
  \includegraphics[scale=0.27]{Images/Average_steps/e.png}
  \includegraphics[scale=0.27]{Images/Average_steps/f.png}
  \centering
  \includegraphics[scale=0.27]{Images/Average_steps/g.png}
  \includegraphics[scale=0.27]{Images/Average_steps/h.png}
\end{subfigure}
\caption{Average number of steps for the agent to reach the goal when trained with rewards generated from brainwaves from OHCs 1-8. Y axis show the averaged number of steps for a run session, while x axis show the number of game matches used to cumulative train the Q-Table.}
\label{fig:avg_steps}
\end{figure}

@@

\begin{figure}[h!]
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.09]{revisedimages/roc_1.eps}   \hspace{-1.5em}
\includegraphics[scale=0.09]{revisedimages/roc_2.eps} \\
%\vspace{-1.3em}
\centering
\includegraphics[scale=0.09]{revisedimages/roc_3.eps}   \hspace{-1.5em}
 \includegraphics[scale=0.09]{revisedimages/roc_4.eps} \\
 \centering
  \includegraphics[scale=0.09]{revisedimages/roc_5.eps}   \hspace{-1.5em}
  \includegraphics[scale=0.09]{revisedimages/roc_6.eps} \\
  \centering
  \includegraphics[scale=0.09]{revisedimages/roc_7.eps}    \hspace{-1.5em}
  \includegraphics[scale=0.09]{revisedimages/roc_8.eps} \\
\end{subfigure}
\caption{ROC Curves for OHCs 1-8. True positive rate is on the vertical axis and false positive rate on the horizontal axis.  The ROC curves for ErrP identification for Subjects 5 and 6 show low classification scores.}
\label{fig:rocsubjects}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{Images/Average_steps/noise.eps}
\caption{Average number of steps for the agent to reach the goal when trained with a classifier produced from sham EEG signals. X axis show the number of gaming agent training matches used to train the Q-Table.}
\label{fig:avg_steps_noise}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\includegraphics[scale=0.60]{revisedimages/transfer_learning_heatmap.eps}
\end{subfigure}
\caption{Heatmap for the transfer learning experiment. Values represent the reduction in the average number of steps required to reach the goal.  Negative values represent net improvements.}
\label{fig:transferlearning}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{Images/Average_steps/all.eps}
\caption{Average steps using Q-Table trained with brainwave session matches from six different OHCs. X axis show the progressive number of gaming agent training matches used to train the Q-Table. These matches correspond to all subjects, excluding subjects 5 and 6 which individually show no significant learning progress.}
\label{fig:avg_steps_all}
\end{figure}

@@

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the variability of the ErrP response in terms of different experiments, levels of 90\%  accuracy in identifying it are reported~\cite{Chavarriaga2014}.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally,  classification results show a low percentage of false positives (Figure~\ref{fig:confusionmatrix}), entailing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.14]{revisedimages/matrix_1.eps}
\includegraphics[scale=0.14]{revisedimages/matrix_2.eps}\\
\includegraphics[scale=0.14]{revisedimages/matrix_3.eps}
\includegraphics[scale=0.14]{revisedimages/matrix_4.eps}\\
\includegraphics[scale=0.14]{revisedimages/matrix_5.eps}
\includegraphics[scale=0.14]{revisedimages/matrix_6.eps}\\
\includegraphics[scale=0.14]{revisedimages/matrix_7.eps}
\includegraphics[scale=0.14]{revisedimages/matrix_8.eps}\\
\end{subfigure}
\caption{Confusion Matrix for OHCs 1-8. Darker colors show higher values.  It can be seen the lower percentage of false positives (upper right corner of each chart).}
\label{fig:confusionmatrix}
\end{figure}

>>>>

% You must have at least 2 lines in the paragraph with the drop letter
% (should never be an issue)
\IEEEPARstart{T}{he} effectiveness of today's human–machine interaction and artificial intelligence is limited by a communication bottleneck, as humans are required to translate high-level concepts into a machine-mandated sequence of instructions~\cite{Xu2020,CURSOR-CONTROL-PAPER}.   Hence, new interaction methods are required to increase the communication bandwidth between computers and humans or to produce alternative communications systems to increase the efficiency of this channel.  In this respect, video games have been widely used as test tools to assess new means of interactions~\cite{Carter2014,Frey2014,Barr2007}. Video gaming agents are computer programs that can sense the computer game environment, process information, and react accordingly within the environment.  They are used in the context of testing and evaluating artificial intelligence algorithms that aim to win the game or to behave like a real user player~\cite{Zhao2020}.
In this work, the feedback obtained from an observational human critic (OHC) in the form of electroencephalographic (EEG) signals is used to evaluate the operational performance of a gaming agent.  Observational human critics are silent subjects observing a computer gaming agent playing the game.

% Reescribí la primera oración porque el sujeto no tenía sentido con el predicado.  TODO
The feasibility of a distinct non-biological communication channel between the Central Nervous System (CNS) and a computer device has been previously proven with Brain Computer Interfaces (BCI) or Brain Machine Interfaces (BMI).~\cite{Vasiljevic2020}.  BCI systems provide a new input modality that can be used in computer games~\cite{Scherer2012,Marshall2013,Nijholt2007}. This advancement is relevant in the context of the accessibility for video games~\cite{Aguado-Delgado2020} and the growing area of e-sports~\cite{Yakovlev2020}.

@@

Given the scenario, Reinforcement Learning (RL)~\cite{Sutton2018} stands out as a natural method to train the agent.  Reinforcement Learning refers to an algorithmic learning strategy inspired on how biological agents learn by exploring their environment while getting negative or positive feedback rewards.  The method aims to maximize positive rewards while minimizing negative feedback.  Thus, the learning problem is posed as a stochastic optimization strategy~\cite{Santos1999}.  Recently, this technique has been used extensively in artificial intelligence~\cite{Nguyen2020}. The influence of DeepBrain's AlphaGo project cannot be neglected, since it was the first to reach a very high proficiency when it won the complex game Go against several world champions~\cite{ALPHA-GO}.

@@

This work unfolds as follows: in Section \ref{section:materials} the general layout of the cognitive game is described. Sections \ref{brainwavesession} and \ref{cognitive_experiment_details} outline the cognitive game procedure used to obtain rewards in the form of ErrP components. Section \ref{q_learning_step_alg} describes the gaming agent learning procedure.  Lastly, results and conclusions are exposed in Sections \ref{results} and \ref{conclusions}.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{images/figure1.png}
    \caption{Overview of the experimental procedure. Brainwaves are obtained by the OpenVibe Acquisition Server.  The Game Manager is responsible for generating the game screen, the game mechanics, and the game movements performed by the gaming agent.  It is also connected to the Acquisition Server to send stimulus information.  The captured information is stored by the OpenVibe Designer.  Offline, EEG signals are classified and they are linked to each game movement calculated by the Game Manager to determine proper rewards for each action.  This information is used by a Reinforcement Learning algorithm that iteratively trains a Q-Table in order to improve the performance of the agent that plays the game.}
    \label{diag:complete_flow}
\end{figure*}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.2]{images/figure2a.png}
\includegraphics[scale=0.2]{images/figure2b.png}
\end{subfigure}
\centering
\caption{Grid system representation used in the cognitive game. The blue spot represents the initial location while the green spot represents the target location. Once the agent reaches the target spot, its color turns red to indicate the end of the play.}
\label{fig:game_representation}
\end{figure}

@@

This phase is divided into a sequence of run sessions and gaming agent training matches.  During the run session, the agent plays 200 matches guided by a specific Q-Table with a 5\% chance of randomly selecting a movement, to reduce deadlocks and loops.  Following the run session, the agent performs a single gaming agent training match.  The gaming agent starts first with a Q-Table initialized with zeros, so the initial policy for the agent is randomized.  For the agent to learn from the feedback generated by the OHC, movement actions are determined by the replay of the agent's actions that were taken during one brainwave session match, in an offline reinforcement learning scheme~\cite{Levine2020}. This allows the Q-Table to be learned based on the OHC's feedback from the movements the agent took, which were executed pseudo-randomly during the brainwave session.  The previously mentioned feedback is not explicit as it comes from the interpreted brain signal data. This implies that the reward is determined by the OHC's brain activity.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{images/figure3a.eps}
    \includegraphics[width=0.5\textwidth]{images/figure3b.eps}
    \includegraphics[scale=1.0]{images/figure3c.png}
    \caption{Grand average of 2-seconds time-locked segments for all OHCs for the "move closer" (top) and "move further" (middle) condition. Zero time on x-axis corresponds to the onset of the stimulus, i.e. when the gaming agent moves. The increased height of the first peak around 0.4 seconds reflects the ErrP response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:grandaverage}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth,height=70pt]{images/figure4.eps}
    %\includegraphics[width=290pt,height=90pt]{revisedimages/avg_5_segments.eps}
    \caption{Binary single trial classification score using five different classifiers while recognizing ErrP potentials for the eight OHCs.   Chance Level is 0.5.}
    \label{fig:classifiers}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.27]{images/figure5a.png}
\includegraphics[scale=0.27]{images/figure5b.png}
\centering
\includegraphics[scale=0.27]{images/figure5c.png}
 \includegraphics[scale=0.27]{images/figure5d.png}
 \centering
  \includegraphics[scale=0.27]{images/figure5e.png}
  \includegraphics[scale=0.27]{images/figure5f.png}
  \centering
  \includegraphics[scale=0.27]{images/figure5g.png}
  \includegraphics[scale=0.27]{images/figure5h.png}
\end{subfigure}
\caption{Average number of steps for the agent to reach the goal when trained with rewards generated from brainwaves from OHCs 1-8. Y axis show the averaged number of steps for a run session, while x axis show the number of game matches used to cumulative train the Q-Table.}
\label{fig:avg_steps}
\end{figure}

@@

\begin{figure}[h!]
\captionsetup[subfigure]{justification=centering}
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.09]{images/figure6a.eps}   \hspace{-1.5em}
\includegraphics[scale=0.09]{images/figure6b.eps} \\
%\vspace{-1.3em}
\centering
\includegraphics[scale=0.09]{images/figure6c.eps}   \hspace{-1.5em}
 \includegraphics[scale=0.09]{images/figure6d.eps} \\
 \centering
  \includegraphics[scale=0.09]{images/figure6e.eps}   \hspace{-1.5em}
  \includegraphics[scale=0.09]{images/figure6f.eps} \\
  \centering
  \includegraphics[scale=0.09]{images/figure6g.eps}    \hspace{-1.5em}
  \includegraphics[scale=0.09]{images/figure6h.eps} \\
\end{subfigure}
\caption{ROC Curves for OHCs 1-8. True positive rate is on the vertical axis and false positive rate on the horizontal axis.  The ROC curves for ErrP identification for Subjects 5 and 6 show low classification scores.}
\label{fig:rocsubjects}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{images/figure7.eps}
\caption{Average number of steps for the agent to reach the goal when trained with a classifier produced from sham EEG signals. X axis show the number of gaming agent training matches used to train the Q-Table.}
\label{fig:avg_steps_noise}
\end{figure}

@@

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\includegraphics[scale=0.60]{images/figure8.eps}
\end{subfigure}
\caption{Heatmap for the transfer learning experiment. Values represent the reduction in the average number of steps required to reach the goal.  Negative values represent net improvements.}
\label{fig:transferlearning}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[scale=0.4]{images/figure9.eps}
\caption{Average steps using Q-Table trained with brainwave session matches from six different OHCs. X axis show the progressive number of gaming agent training matches used to train the Q-Table. These matches correspond to all subjects, excluding subjects 5 and 6 which individually show no significant learning progress.}
\label{fig:avg_steps_all}
\end{figure}

@@

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, considering the variability of the ErrP response in terms of different cognitive experiments, levels of 90\%  accuracy in identifying it, are reported~\cite{Chavarriaga2014}.  In this work, despite the low accuracy in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally,  classification results show a low percentage of false positives (Figure~\ref{fig:confusionmatrix}), entailing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.

\begin{figure}[h!]
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[scale=0.14]{images/figure10a.eps}
\includegraphics[scale=0.14]{images/figure10b.eps}\\
\includegraphics[scale=0.14]{images/figure10c.eps}
\includegraphics[scale=0.14]{images/figure10d.eps}\\
\includegraphics[scale=0.14]{images/figure10e.eps}
\includegraphics[scale=0.14]{images/figure10f.eps}\\
\includegraphics[scale=0.14]{images/figure10g.eps}
\includegraphics[scale=0.14]{images/figure10h.eps}\\
\end{subfigure}
\caption{Confusion Matrix for OHCs 1-8. Darker colors show higher values.  It can be seen the lower percentage of false positives (upper right corner of each chart).}
\label{fig:confusionmatrix}
\end{figure}
